\LARGE \bf Towards a Data-driven Approach to Motion Planning

### 2014-10-01 paper2015asml.tex
learning part + cleanup of features

\begin{figure}
\centering
  \includegraphics[trim = 70mm 0mm 70mm 0mm, width=0.62\columnwidth, clip=true]{pictures/countertop_scene2}
  \begin{minipage}[b][][s]{0.352\columnwidth}
  \centering
    \includegraphics[trim = 70mm 0mm 70mm 0mm, width=\textwidth, clip=true]{pictures/drawer_scene2}
    \vfill
    \includegraphics[trim = 30mm 0mm 30mm 0mm, width=\textwidth, clip=true]{pictures/industrial_scene2}
  \end{minipage}
\caption{The scenes used in generating the library of robot motions to collect feedeback for. \emph{Drawer} (top right), \emph{Countertop} (left), and \emph{Industrial} (bottom right)}\label{fig:scenes}
\end{figure}

@@

\begin{table*}
\centering
\begin{tabular}{|c||c|c|c|c|c|c|c|}
\hline
Efficiency  & Highly inefficient& Moderately inefficient& Slightly inefficient& Neutral & Slightly efficient& Moderately efficient& Highly efficient\\
\hline
Elegance    & Highly awkward& Moderately awkward& Slightly awkward& Neutral & Slightly elegant& Moderately elegant& Highly elegant\\
\hline
Smoothness  & Highly rough& Moderately rough& Slightly rough& Neutral & Slightly smooth& Moderately smooth& Highly smooth\\
\hline
Overall     & Extremely poor& Moderately poor& Somewhat poor& Neutral & Somewhat good& Moderately good& Extremely good\\
\hline
\end{tabular}
\caption{The rating categories and the associated scale presented to the survey participant}
\label{tab:rating_scales}
\end{table*}

Robot motion study for industrial has been primarily influenced by the need to generate smooth jerk-free trajectories that minimize wear and tear of internal parts. In recent years, there has been more interest in the generation of motions for robots working in human environments, i.e. environments with clutter and people working next to the robots. Most attention, though, has been focused on generating {\em collision-free} motions. Examples include the use of randomized planners~\cite{kuffner2000rrt}, trajectory optimization techniques~\cite{Mrinal:2011}, and deterministic search-based techniques~\cite{cohen2010search}. Such approaches have also been modified to account for other types of constraints, e.g. dynamic constraints or process constraints arising from the task.

@@

The MoveIt! benchmarking architecture~\cite{cohen2012generic} was extended with video recording capabilities to quickly generate a large number of trajectories. The MoveIt! architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries). It also allows for the use of any supported planning algorithms (planners). A combination of different scenes, motion queries, and planners was used to generate robot trajectories for evalutation. We chose three distinct scenes (Fig.~\ref{fig:scenes}): a kitchen ({\sc Countertop}), an industrial workbench ({\sc Industrial}) and a drawer on a table ({\sc Drawer}). Multiple queries were generated in each scene for motions of the right arm of the PR2 robot. The robot was positioned in the scenes manually and motion planning queries were constructed by hand. The positioning in the scenes was done with the expectation that the planning queries would not be impossible or trivial for the planning algorithms to solve.

@@

The planning algorithms that we used are RRT-Connect~\cite{kuffner2000rrt} and RRT*.  RRT*~\cite{frazzoli-RRTstar} was used with two different cost functions: {\sc PathLength} and {\sc MaxMinClearance}.The objective for {\sc PathLength} minimizes the length of the trajectory (as the sum of the euclidean distances between ends of trajectory segments). The objective for {\sc MaxMinClearance} computes uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize. Both planners are implemented in the OMPL library~\ref{OMPL:2012}.

@@

The infrastructure we designed around MoveIt! allows us to automatically generate movies of the generated trajectories using the ROS Visualizer (Rviz). We generated movies from different viewpoints to maximize the visibility of the motion to human observers reviewing it. Each participant was allowed to provide feedback no more than once for each of the trajectories generated by the survey design. We ran the survey for 103 trajectories, requesting no more than 20 ratings per trajectory. Participants are allowed to rate multiple distinct trajectories, but no more than once. The analysis of these results is presented in Section~\ref{sec:analysis}.

@@

We conducted an online survey{\footnote The online survey was conducted after gaining a waiver from SRI's IRB. Our survey followed standard guidelines for privacy of participant information when conducting a survey on Mechanical Turk.}  to ask human observers to rate a variety of trajectories of the PR2 robot. The surveys were built on Amazon's Mechanical Turk~\cite{paolacci2010running}, which has become a useful tool for large online data gathering and crowd-sourced work. Each respondent in the survey was presented with a set of demographic questions, a video, and four scales to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression. Each respondent was expected to rate multiple trajectories but could not rate a trajectory more than once. Each video has four different views of the robot displayed simultaneously. Participants were asked a demographic questions about (a) familiarity with video games, (b) familiarity with robotics, (c) Age group: categorized as 18-24, 25-44, 45-64 and 65+ and (d) Highest Education Level: PhD/Masters/Bachelors/Associate/High School. None of these pieces of information were used to filter the responses we collected.  A summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}.

Mechanical Turk allows monetary compensation on completion of the survey, which we set to 15\textcent~ per survey. We estimated the time that each participant would take to rate a single trajectory as 1-2 minutes. Additionally, the participants were allowed to replay the trajectory video repeatedly. The maximum possible payout for each participant is thus 15.45\textdollar (if each survey took 1.5 minutes to complete then the wage is approximately 6\textdollar~ per hour). This is standard compensation for quality work on Mechanical Turk for tasks of this type, as was pointed out by several veteran Mechanical Turk surveyors.

Respondents on Mechanical Turk are only paid if the survey organizer approves their response. A control question, asking the respondent to identify the known action of a robot in a video, served as the first satisficing trap to ensure that participants were truthfully attempting the survey (see~\cite{krosnick1991response}~\cite{krosnick1996satisficing} for more information on satisficing in survey responses. Survey responses were also manually examined to assess whether participants were responding honestly. We looked at whether participant responses were outliers when compared to the rest of the group. We also examined the average time taken by each participant to complete the survey. Overall, we found that all the participants had made a honest attempt at filling out the rating and we did not reject any responses.

%\begin{figure}
%    \frame{\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/amazon_survey_screenshot}}
%    \caption{The main survey question which prompts the participant for a rating in four motion characteristics (efficiency, elegance, smoothness and overall impression) after presenting a video that demonstrates the motion from four clearly visible vantage points for any sample trajectory.}
%    \label{fig:survey_question}
%\end{figure}

@@

\subsubsection{Survey Design and Process}

%\begin{figure}
%\frame{\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/amazon_control_question}}
%\caption{The survey control question which prompts the participant about the action of the robot in the accompanying video. The question we adopted was to ask the particpant which colored cube the robot touched.}
%\label{fig:control_question}
%\end{figure}

\section{Survey Results}
\label{sec:analysis}
The results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of (average) responses for each of the four survey questions across the 103 trajectories that were rated by users. It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well. An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall). Table~\ref{table:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes. Attributes across individual questions are highly correlated at a statistically significant level of 1% (with p-values < 0.01).

%\begin{table}
%\label{table:correlation}
%\begin{tabular}
%\end{tabular}
%\end{table}

Confidence intervals for the overall rating of each trajectory are shown in Fig.~\ref{confidence}. The small spread in the confidence intervals clearly indicates that different users had similar opinions about the same trajectory. Indeed, we found little evidence of any personal preference or clustering of users based on their preference for certain trajectories over others.

\begin{figure}[t]
  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
  \caption{Histogram of averaged ratings for the 103 different trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Inelegant, Smooth vs. Rough, Overall Rating}
\label{fig:survey_raw}
\end{figure}

\section{Feature Design for Trajectory Classification}

The trajectory features we selected for the classification task consider the environment, the shape of the trajectory and the dynamics-related quantities. These make up our unary features computed only on the input trajectory.

We also include features that compare each trajectory to a nominal trajectory for its problem which also serve to characterize how much the trajectory deviates from the "expected" path. We go over the selection of the nominal path in Section~\ref{subsec:multi_traj_feat}.

\subsection{Unary Trajectory Features}

Here we detail the features we computed on the trajectory \tj{} with waypoints $q_i$ where $i={0,\mathellipsis,n}$ and $n=|\tj{}|$.

\subsubsection{Path length}

The length of the trajectory \tj{} computed as the sum over trajectory segments of the euclidean distance between end points of the segment, or

\begin{equation}
\text{Length} = \sum_{i=0}^{n-1} \| q_\text{i}-q_\text{i+1} \|
\end{equation}

We compute Path Length for the joint space trajectory and for three workspace trajectories that are dervied from the forward kinematic map of the joint-space trajectory for the following links in the kinematic chain: for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$).

\subsubsection{Clearance}

The average clearance of a trajectory \tj{} is the average minimum distance of configurations at waypoints to the closest obstacle, or

\begin{equation}
\text{Clearance} = \frac{1}{n} \sum_{i=0}^n \min_{o \in O} \text{dist}(q_i, o)
\end{equation}

where $O$ represents the set of obstacles in the scene and $\text{dist}(q_i, o)$ calculates the closest distance between the kinematic bodies of obstacle $o$ and configuration $q_i$.

\subsubsection{Smoothness}

The smoothness of a trajectory is computed by averaging the included angle of three consecutive waypoint configurations, for all triplets of waypoints in \tj{}. The angle is computed using joint-space euclidean distance between configurations as the length of sides of the triangle.

\begin{equation}
\begin{aligned}
&\text{Smoothness} = \frac{1}{n-2} \sum_{i=1}^{n-1} \mathellipsis\\
&\left[ \pi- \text{angle}(\|q_\text{i-1}-q_\text{i}\|,\|q_\text{i}-q_\text{i+1}\|,\|q_\text{i+1}-q_\text{i-1}\|) \right]\\
\end{aligned}
\end{equation}

where the included angle is computed

\begin{equation}
\text{angle}(a,b,c) = \arccos (a^2+c^2-b^2)/2ac
\end{equation}

We compute smoothness for the joint-space trajectory \tj{} and for workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.

\subsubsection{Maximum Acceleration}

For a joint-space trajectory \tj{} consisting of waypoints, the maximum acceleration is computed by finding the accelerations at the knot points of splines fit to each joint's trajectory. The maximum over waypoints of the norm of these accelerations is the feature

\begin{equation}
\text{Max Acceleration} = \max_{i={0,\mathellipsis,n}} {\ddot{q}_i}^T\ddot{q}_i
\end{equation}

We compute maximum acceleration for the joint-space trajectory \tj{} and for workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.

\subsubsection{Maximum Radius of Curvature}

Curvature at a given waypoint is computed by first fitting the work-space trajectory with splines for each coordinate and computing from the higher order derivatives

\begin{equation}
R_i = \frac{\|\dot{q}_i\|^3}{\|\dot{q}_i \times \ddot{q}_i\|}
\end{equation}

where velocity $\dot{q}_i$ and acceleration $\ddot{q}_i$ are computed at the waypoint $i$ from differentiating the interpolating spline of the work-space trajectory. Alternatively, the higher order derviatives can be computed using finite differencing methods.

We compute maximum curvature for only the workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.

\subsection{Multi-Trajectory Features}
\label{subsec:multi_traj_feat}

We compared the trajectories to nominal end-effector trajectories that were created using a breadth-first search in a 3D voxel approximation of the planning problem.

The planning scene is discretized, and voxels are marked as occupied or empty if any part of the voxel is considered in collision with the underlying planning scene.

The start and goal voxel is simply the $(x,y,z)$ of the end-effector tool-frame. The free voxels then make up our vertices $V_\text{grid}$ for our graph.

The edges $E_\text{grid}$ used in this search are the 26-connected grid of adjacent neighboring voxels. For example, some valid adjacent voxels can be computed by adding $(\Delta x,\Delta y,\Delta z)$ to the current coordinates, where $\Delta x= \Delta y= \Delta z= \{\pm1,0\}$. An edge only exists in this graph if the ending voxel is collision.

The order the neighboring voxels are generated by the breadth-first search is done such that neighboring voxels that share \emph{faces} are pushed onto the queue before those that share an \emph{edge}, which are then added before those that share a \emph{corner} with the generating state. This lets us use a simple BFS to search the graph with unit cost edges.

Using the graph $G=(V_\text{grid}, E_\text{grid})$, we perform breadth-first search to get a collision free path from $V_\text{start}$ to $V_\text{end}$. The resulting path is given a nominal timing and we refer to it as $\xi_{\text{BFS}}$.

We compute comparison features between $\xi_{\text{BFS}}$ and three other workspace trajectories that are computed from the forward kinematic map of the joint-space trajectory for various links in the kinematic chain, which are the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$).

We compared them by computing the the Hausdorff distances~\cite{dubuisson1994modified}, and Dynamic Time Warping~\cite{senin2008dynamic} costs between $\xi_{\text{BFS}}$ and each of the other three.

@@

Dynamic Time Warping was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}\cite{chen2013dynamic}. It provides a way to compute a distance despite differences in timing of the trajectories as well.

DTW computes a mapping of a subset of the waypoints in $\xi_\text{A}$ to a subset of the waypoints in $\xi_\text{B}$ which constitutes the warping path, or

\begin{equation}
\xi_\text{DTW} = \{ (0, 0), (a_i, b_j), \mathellipsis ,(n_A, n_B) \}
\end{equation}

where $n_A = |\tj{A}|$ and $n_B = |\tj{B}|$ which are the number of waypoints in the trajectory.

@@

>>>>

\begin{figure}
\centering
  \includegraphics[trim = 70mm 0mm 70mm 0mm, width=0.62\columnwidth, clip=true]{pictures/countertop_scene2}
  \begin{minipage}[b][][s]{0.352\columnwidth}
  \centering
    \includegraphics[trim = 70mm 0mm 70mm 0mm, width=\textwidth, clip=true]{pictures/drawer_scene2}
    \vfill
    \includegraphics[trim = 30mm 0mm 30mm 0mm, width=\textwidth, clip=true]{pictures/industrial_scene2}
  \end{minipage}
\caption{The scenes used in generating the robot motions to collect feedback from human observers. \emph{Drawer} (top right), \emph{Countertop} (left), and \emph{Industrial} (bottom right)}\label{fig:scenes}
\end{figure}

@@

\begin{table*}
\centering
\begin{tabular}{|c||c|c|c|c|c|c|c|}
\hline
Efficiency  & Highly inefficient& Moderately inefficient& Slightly inefficient& Neutral & Slightly efficient& Moderately efficient& Highly efficient\\
\hline
Elegance    & Highly awkward& Moderately awkward& Slightly awkward& Neutral & Slightly elegant& Moderately elegant& Highly elegant\\
\hline
Smoothness  & Highly rough& Moderately rough& Slightly rough& Neutral & Slightly smooth& Moderately smooth& Highly smooth\\
\hline
Overall     & Extremely poor& Moderately poor& Somewhat poor& Neutral & Somewhat good& Moderately good& Extremely good\\
\hline
\end{tabular}
\caption{The rating categories and the associated scale presented to the survey respondent}
\label{tab:rating_scales}
\end{table*}

Robot motion for industrial applications has been primarily influenced by the need to generate smooth jerk-free trajectories that minimize wear and tear of internal parts. In recent years, there has been more interest in the generation of motions for robots working in human environments, i.e. environments with clutter and people working next to the robots. Most attention, though, has been focused on generating {\em collision-free} motions. Examples include the use of randomized planners~\cite{kuffner2000rrt}, trajectory optimization techniques~\cite{Mrinal:2011}, and deterministic search-based techniques~\cite{cohen2010search}. Such approaches have also been modified to account for other types of constraints, e.g. dynamic constraints or process constraints arising from the task.

@@

The MoveIt! benchmarking architecture~\cite{cohen2012generic} was extended with video recording capabilities to quickly generate a large number of trajectories. The MoveIt! architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries). It also allows for the use of any supported planning algorithms (planners). A combination of different scenes, motion queries, and planners was used to generate robot trajectories for evalutation. We chose three distinct scenes (Fig.~\ref{fig:scenes}): a kitchen ({\sc Countertop}), an industrial workbench ({\sc Industrial}) and a drawer on a table ({\sc Drawer}). Multiple queries were generated in each scene for trajectories of the right arm of the PR2 robot. The robot was positioned in the scenes manually and motion planning queries were constructed by hand. The positioning in the scenes was done with the expectation that the planning queries would not be impossible or trivial for the planning algorithms to solve.

@@

The planning algorithms that we used are RRT-Connect~\cite{kuffner2000rrt} and RRT*.  RRT*~\cite{frazzoli-RRTstar} was used with two different cost functions: {\sc PathLength} and {\sc MaxMinClearance}.The objective for {\sc PathLength} minimizes the length of the trajectory (as the sum of the euclidean distances between ends of trajectory segments). The objective for {\sc MaxMinClearance} computes uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize. Both planners are implemented in the OMPL library~\ref{OMPL:2012}. The infrastructure we designed around MoveIt! allows us to automatically generate movies of the generated trajectories using the ROS Visualizer (Rviz). We generated movies from different viewpoints to maximize the visibility of the motion to human observers reviewing it. A total of 103 different trajectories were generated in this manner.

@@

@@

We conducted an online survey\footnote{The online survey was conducted after gaining a waiver from SRI's IRB. Our survey followed standard guidelines for privacy of respondent information when conducting a survey on Mechanical Turk.} where we asked human observers to rate a variety of trajectories of the PR2 robot. The surveys were built on Amazon's Mechanical Turk~\cite{paolacci2010running}, which has become a useful tool for large online data gathering and crowd-sourced work. Each respondent in the survey was presented with a set of demographic questions, a video, and four scales to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression. Each respondent was expected to rate multiple trajectories but could not rate a trajectory more than once. Each video has four different views of the robot displayed simultaneously. respondents were asked demographic questions about familiarity with video games, familiarity with robotics, age group and highest education level. None of these pieces of information were used to filter the responses we collected.  A summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}.

Mechanical Turk allows monetary compensation on completion of the survey, which we set to 15\textcent~ per survey. We estimated the time that each respondent would take to rate a single trajectory as 1-2 minutes. Additionally, the respondents were allowed to replay the trajectory video repeatedly. The maximum possible payout for each respondent is thus 15.45\textdollar (if each survey took 1.5 minutes to complete then the wage is approximately 6\textdollar~ per hour). This is standard compensation for quality work on Mechanical Turk for tasks of this type, as was pointed out to us by several veteran Mechanical Turk surveyors.

Respondents on Mechanical Turk are only paid if the survey organizer approves their response. A control question, asking the respondent to identify the known action of a robot in a video, served as the first satisficing trap to ensure that respondents were truthfully attempting the survey (see~\cite{krosnick1991response}~\cite{krosnick1996satisficing} for more information on satisficing in survey responses. Survey responses were also manually examined to assess whether respondents were responding honestly. We looked at whether respondent responses were outliers when compared to the rest of the group. We also examined the average time taken by each respondent to complete the survey. Overall, we found that all the respondents had made a honest attempt at filling out the rating and we did not reject any responses.

%\begin{figure}
%    \frame{\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/amazon_survey_screenshot}}
%    \caption{The main survey question which prompts the respondent for a rating in four motion characteristics (efficiency, elegance, smoothness and overall impression) after presenting a video that demonstrates the motion from four clearly visible vantage points for any sample trajectory.}
%    \label{fig:survey_question}
%\end{figure}

@@

%\begin{figure}
%\frame{\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/amazon_control_question}}
%\caption{The survey control question which prompts the respondent about the action of the robot in the accompanying video. The question we adopted was to ask the particpant which colored cube the robot touched.}
%\label{fig:control_question}
%\end{figure}

\section{Survey Results and Analysis}
\label{sec:analysis}
The survey resulted in at least 20 ratings for each trajectory. While the raw results are interesting in themselves, we are most interested in knowing why the respondents preferred particular trajectories. We aim to infer this information using a data-driven approach. We compute multiple features for each trajectory and use the features and survey ratings to design a classifier that can classify a new trajectory into one of three classes: low human-observer preference, medium human-observer preference and high human-observer preference. The classifier can thus be used with new trajectories to predict the expected rating of that trajectory by a human-bserver. The classifier also helps to identify the features that are most key to determining the human-observer preference for a particular trajectory over others. These features will be key in eventually developing new motion planners or cost functions for motion planners that create plans that are more acceptable to human observers.

\subsection{Raw Survey Results}
\begin{figure}[t]
  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
  \caption{Histogram of averaged ratings for the 103 different trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Inelegant, Smooth vs. Rough, Overall Rating}
\label{fig:survey_raw}
\end{figure}

The raw results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of (average) responses for each of the four survey questions across the 103 trajectories that were rated by users. It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well. An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall). Table~\ref{table:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes. Attributes across individual questions are highly correlated at a statistically significant level of 1% (with p-values < 0.01). Confidence intervals for the overall rating of each trajectory are shown in Fig.~\ref{confidence}. The small spread in the confidence intervals clearly indicates that different users had similar opinions about the same trajectory.

\begin{table}
\label{table:correlation}
\begin{centering}
\begin{tabular}{|c|c|c|c|c|}
\hline
& Efficient & Elegant & Smooth & Overall \\ \hline
Efficient & 1 & 0.96 & 0.91 & 0.98 \\ \hline
Elegant & 0.96 & 1 & 0.96 & 0.98 \\ \hline
Smooth & 0.91 & 0.96 & 1 & 0.96 \\ \hline
Overall & 0.98 & 0.98 & 0.96 & 1 \\ \hline
\end{tabular}
\end{centering}
\caption{Correlation between the four attributes in the survey.}
\end{table}

\subsection{Trajectory Features}

The first set of features we compute for each trajectory take into account the environment around the robot, the shape of the trajectory and the dynamics of the trajectory. These make up our unary features computed only on the input trajectory. We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the "expected" path. We go over the selection of the nominal path in Section~\ref{subsec:multi_traj_feat}. First, we detail the features we computed on each trajectory \tj{} with waypoints $q_i$ where $i={0,\mathellipsis,n}$ and $n=|\tj{}|$. Note that our features are generic, i.e. they can be computed in the same manner for any robot arm trajectory. We believe that our methods should be easily extensible to robots other than the PR2 as well and individual trajectory classifiers could be built for each different robot arm in a similar manner.

{\bf Path length:}
%The length of the trajectory \tj{} computed as the sum over trajectory segments of the euclidean distance between end points of the segment, or
%\begin{equation}
%\text{Length} = \sum_{i=0}^{n-1} \| q_\text{i}-q_\text{i+1} \|
%\end{equation}
We compute path length for the joint-space trajectory and for three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links of the right arm (the moving arm).

{\bf Clearance:}
The average clearance of a trajectory \tj{} is the average minimum distance of configurations at waypoints to the closest obstacle, or
\begin{equation}
\text{Clearance} = \frac{1}{n} \sum_{i=0}^n \min_{o \in O} \text{dist}(q_i, o)
\end{equation}
where $O$ represents the set of obstacles in the scene and $\text{dist}(q_i, o)$ calculates the closest distance between obstacle $o$ and configuration $q_i$.

{\bf Smoothness:}
The smoothness of a trajectory is computed by averaging the included angle of three consecutive waypoint configurations, for all triplets of waypoints in \tj{}. The angle is computed using joint-space euclidean distance between configurations as the length of sides of the triangle.
\begin{equation}
\begin{aligned}
&\text{Smoothness} = \frac{1}{n-2} \sum_{i=1}^{n-1} \mathellipsis\\
&\left[ \pi- \text{angle}(\|q_\text{i-1}-q_\text{i}\|,\|q_\text{i}-q_\text{i+1}\|,\|q_\text{i+1}-q_\text{i-1}\|) \right]\\
\end{aligned}
\end{equation}
where the included angle is computed
\begin{equation}
\text{angle}(a,b,c) = \arccos (a^2+c^2-b^2)/2ac
\end{equation}
We compute smoothness for both the joint-space trajectory \tj{} and for workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.

{\bf Maximum Acceleration:}
We compute the absolute maximum acceleration for the joint-space trajectory \tj{} and for workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.
%For a joint-space trajectory \tj{} consisting of waypoints, the maximum acceleration is computed by finding the accelerations at the knot points of splines fit to each joint's trajectory. 
%\begin{equation}
%\text{Max Acceleration} = \max_{i={0,\mathellipsis,n}} {\ddot{q}_i}^T\ddot{q}_i
%\end{equation}

{\bf Maximum Radius of Curvature: }
Curvature at a given waypoint is computed by splines to the workspace trajectory. We compute maximum curvature for only the workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.
% for each coordinate and computing from the higher order derivatives
%\begin{equation}
%R_i = \frac{\|\dot{q}_i\|^3}{\|\dot{q}_i \times \ddot{q}_i\|}
%\end{equation}
%where velocity $\dot{q}_i$ and acceleration $\ddot{q}_i$ are computed at the waypoint $i$ from differentiating the interpolating spline of the work-space trajectory. Alternatively, the higher order derviatives can be %computed using finite differencing methods.

\subsection{Multi-Trajectory Features}
\label{subsec:multi_traj_feat}
We wanted to design a feature that captured the difference between the planned trajectory and an expected path for the end-effector in the environment. The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D voxel approximation of the planning problem. The environment model was first discretized into a voxel-grid, and voxels are marked as occupied or empty. Any voxel that is partially occupied is considered to be occupied. A 26-connected grid was used in the planning process to find a path in the voxel grid for a cubical shape moving from the start position of the end-effector of the arm to the desired goal position.  Essentially, this would be the optimal path followed by the end-effector from start to goal if it were modeled as a free-standing cube. Next, we compute comparison features between $\xi_{\text{BFS}}$ and three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links. The comparison features computed between each pair of trajectories are the Hausdorff distances~\cite{dubuisson1994modified} and Dynamic Time Warping~\cite{senin2008dynamic} distances.

%The start and goal voxel is simply the $(x,y,z)$ of the end-effector tool-frame. The free voxels then make up our vertices $V_\text{grid}$ for our graph.

%The edges $E_\text{grid}$ used in this search are the 26-connected grid of adjacent neighboring voxels. For example, some valid adjacent voxels can be computed by adding $(\Delta x,\Delta y,\Delta z)$ to the current %coordinates, where $\Delta x= \Delta y= \Delta z= \{\pm1,0\}$. An edge only exists in this graph if the ending voxel is collision.

%The order the neighboring voxels are generated by the breadth-first search is done such that neighboring voxels that share \emph{faces} are pushed onto the queue before those that share an \emph{edge}, which are then %added before those that share a \emph{corner} with the generating state. This lets us use a simple BFS to search the graph with unit cost edges.

%Using the graph $G=(V_\text{grid}, E_\text{grid})$, we perform breadth-first search to get a collision free path from $V_\text{start}$ to $V_\text{end}$. The resulting path is given a nominal timing and we refer to it as %$\xi_{\text{BFS}}$.

@@

Dynamic Time Warping (DTW) was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}\cite{chen2013dynamic}. It provides a way to compute a distance despite differences in timing of the trajectories as well. DTW computes a mapping of a subset of the waypoints in $\xi_\text{A}$ to a subset of the waypoints in $\xi_\text{B}$ which constitutes the warping path, or

\begin{equation}
\xi_\text{DTW} = \{ (0, 0), (a_i, b_j), \mathellipsis ,(n_A, n_B) \}
\end{equation}
where $n_A = |\tj{A}|$ and $n_B = |\tj{B}|$ which are the number of waypoints in the trajectory.

@@

\subsection{Classification}
The 103 data points were broken up into 3 classes based on average overall ratings:
\begin{itemize}
\item “low” - 11 data points $\in (1,3]$,
\item “medium”  - 29 data points $\in (3,5]$,
\item “high” – 63 data points $\in (5,7]$.
\end{itemize}
The data was divided into two samples, a training sample and a test sample, with 1/3rd of each class (overall rating) used to construct the test data sample and the remaining 2/3rd data points within each class are used to train the various models. This results in a training set consisting of 68 data points and test set consisting of 35 data points. Table~\ref{table:test_training} shows how the average overall ratings are distributed across training and test sets.
\begin{table}[h]
\begin{centering}
\label{table:test_training}
\begin{tabular}{|c|c|c|c|}
\hline
 & low (1,3] & medium (3,5] & high (5,7] \\ \hline
Training & 7 & 19 & 42 \\ \hline
Test & 4 & 10 & 21 \\ \hline
\end{tabular}
\caption{Distribution between training and test samples.}
\end{centering}
\end{table}

In all, we utilized three different classification frameworks to predict overall ratings-decision trees, random forests and boosting. As a base framework we first construct a decision tree model.  Decision trees for classification provide a good visual representation especially in the presence of complex features, where the relationships between predictors and response is not immediately obvious. Figure ~\ref{fig:decision_tree} shows our base decision tree model. The features of interest in this tree include the DTW distance computed for the wrist, the hausdorf distance computed for the end-effector and the max curvature computed for the elbow. Using 10-fold cross-validation, we also design a simpler tree that has just three terminal nodes with the $wrist_tool_dtw$ feature serving as the key discriminant among the classes. This tree has an accuracy of 86\% on the test data. It is clear that the BFS path, which serves as the nominal path for computing both the DTW and Hausdorf distance, is a good approximation of the path expected by the human observer for the end-effector of the robot. It is also clear that paths where the elbow has unnecessarily large motions are also penalized by human observers.

The second classification framework we use is a random forest ensemble of trees. This involves randomly dividing the data into several bootstrap training samples and building a tree on each bootstrap sample. Within each boostrapped sample, at each split a subset of features is randomly selected from the full set of features, but only one predictor is used at each split. This brings additional randomness within each tree. The class is chosen by majority vote across all trees. For every tree constructed from a bootstrapped training sample there is one Out of Bag (OOB) error which can be thought of as a test error for every tree. It is estimated internally, during the run. On test data, the random forest method returns 83\% accuracy.

Fig.~\ref{fig:feature_importance_1} and Fig.~\ref{fig:feature_importance_2} show two measures of variable (or feature) importance. The first measure shows the mean decrease in accuracy for each feature. The importance of a feature is determined by how much the accuracy of the random forest is affected by comparing, for each tree, the OOB error with the error obtained after permuting each feature. The differences are averaged across trees and normalized.  The second measure is based on the Gini index and measures the total decrease in node heterogeneity/impurities from splitting on the feature.  For both measures higher values denote that the features are more important.

\begin{figure}[t]
  \centering{\includegraphics[width=0.99\linewidth]{pictures/meandecrease}}
  \caption{Mean decrease in accuracy for each feature.}
\label{fig:feature_importance_1}
\end{figure}

\begin{figure}[t]
 \centering{\includegraphics[width=0.99\linewidth]{pictures/gini}}
  \caption{Gini index (higher values indicate more important features).}
\label{fig:feature_importance_2}
\end{figure}
